Airspace configuration decision based on Monte-Carlo tree search

### 2018-04-21 article/article.tex
(~) Some corrections in model part
path -> \pi, array -> cases, no \times sign

In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
to be defined. This cost depend on the workload in each the sector. We give the
following definitions~\cite{ferrari2017} to represent a high
(respectively normal and low) cost for the partition:
\begin{itemize}
  \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) * p_{high}^{S, t} * |S|^2$
  \item $c_=(P, t) = 1 / (\sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$)
  \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) * p_{low}^{S, t} * |S|^{-2}$
\end{itemize}
with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equal 1
if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
the two others, and 0 otherwise.

It is now possible to assign a cost to each partition. This cost is linear
regarding $c_+$, $c_=$, $c_-$ and $n$ (cardinal of the partition $P$). The
partition cost is defined as follow:
\begin{equation}
C(P, t) = \alpha c_+ + \beta c_= + \gamma c_- +\lambda n  
\end{equation}

The parameters $\alpha$, $\beta$, $\gamma$ and $\lambda$ (all positive)
determine a priority on which parameter to optimize. For instance, a high
value $\alpha$ help to
minimize $c_+$, hence the overall number of sectors with too much tafic.
In a real application, it maybe interesting to order those parmaeters as
follow : $\alpha > \gamma > \beta > \lambda$.

@@

In an operational context, each reconfiguration raise the workload for
controllers. This trend needs to be considered and a transition cost needs to
be defined for this matter. A transition is (TODO), and the transition cost is
then:
\begin{equation}
  C_{tr}(P_1, P_2) =
\left\{
\begin{array}{l}
    0 \qquad if \qquad P_1 = P_2 \\
    1 \qquad otherwise
\end{array}
\right.
\end{equation}

@@

With a cost defined for partition and transition, it is now possible to
aggregate everything in order to build a cost function over an entire
path. For a path ($path = [P_0, ..., P_n]$) with a time from $t_0$ to $t_n$,
the cost function is given by:
\begin{equation}
  \begin{split}
  f(path) = C(P_0, t_0) + \sum_{i = 1}^{n} [C(P_i, t_i) +\\ \theta \times C_{tr}(P_{i-1}, P_i)]   
  \end{split}
\end{equation}

with $\theta > 0$ a parameter to determine.
This is the imediate function to minimize in the A$^*$ algorithm.

The Monte Carlo Tree search algorithm maximize a objective function. This function
represent a potential reward over a path, i.e. a list of partitions over time.
Given the loss function defined previously, the target function can be
construct as:
\begin{equation}
  \psi(path) = \frac{1}{f(path)}
\end{equation}

>>>>

In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
to be defined. This cost depend on the workload in each the sector. We give the
following definitions~\cite{ferrari2017} to represent a high
(respectively normal and low) cost for the partition:
\begin{itemize}
  \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) p_{high}^{S, t} |S|^2$
  \item $c_=(P, t) = 1 / \left(
      \sum_{S \in P_t} \delta_{n}(S, t) p_{norm}^{S} |S|^{-2}
    \right)$
  \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) p_{low}^{S, t} |S|^{-2}$
\end{itemize}
with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equal 1
if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
the two others, and 0 otherwise.

It is now possible to assign a cost to each partition. This cost is linear
regarding $c_+$, $c_=$, $c_-$ and $n$ (cardinal of the partition $P$). The
partition cost is defined as follows:
\begin{equation}
  C(P, t) = \alpha c_+ + \beta c_= + \gamma c_- +\lambda n
\end{equation}

The parameters $\alpha$, $\beta$, $\gamma$ and $\lambda$ (all positive)
determine a priority on which parameter to optimize. For instance, a high
value $\alpha$ help to
minimize $c_+$, hence the overall number of sectors with too much tafic.
In a real application, it maybe interesting to order those parmaeters as
follow: $\alpha > \gamma > \beta > \lambda$.

@@

In an operational context, each reconfiguration raise the workload for
controllers. This trend needs to be considered and a transition cost needs to
be defined for this matter. A transition is (TODO), and the transition cost is
then:
\begin{equation}
  C_{tr}(P_1, P_2) =
  \begin{cases}
    0 & \text{if\,} P_1 = P_2\\
    1 & \text{otherwise}
  \end{cases}
\end{equation}

@@

With a cost defined for partition and transition, it is now possible to
aggregate everything in order to build a cost function over an entire
path. For a path $\pi = [P_0, \dots, P_n]$ with a time from $t_0$ to $t_n$,
the cost function is given by:
\begin{equation}
  \begin{split}
    f(\pi) = C(P_0, t_0) + \sum_{i = 1}^{n} [C(P_i, t_i) +\\
    \theta C_{tr}(P_{i-1}, P_i)]
  \end{split}
\end{equation}

with $\theta > 0$ a parameter to determine.
This is the immediate function to minimize in the A$^*$ algorithm.

The Monte Carlo Tree search algorithm maximize a objective function. This function
represent a potential reward over a path, i.e.\ a list of partitions over time.
Given the loss function defined previously, the target function can be
construct as:
\begin{equation}
  \psi(\pi) = \frac{1}{f(\pi)}
\end{equation}
